### 2017-03-15 paper/main.tex
Started on paper

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr} \usepackage{times} \usepackage{epsfig}
\usepackage{graphicx} \usepackage{amsmath} \usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{1} \begin{document}

%%%%%%%%% TITLE
\title{\LaTeX\ Author Guidelines for CVPR Proceedings}

\author{Weilian Song\\ 
University of Kentucky\\ 
{\tt\small
weilian.song@uky.edu}
% For a paper whose authors are all at the same institution, omit the
% following lines up until the closing ``}''.  Additional authors and
% addresses can be added with ``\and'', just like the second author.  To
% save space, use either the email address or home page, not both
\and Nathan Jacobs\\ 
University of Kentucky\\ 
{\tt\small jacobs@cs.uky.edu}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract} TODO: Make Abstract last \end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

Acoustic detection of birds in the wild is useful in many biological
scenarios, including estimation of bird population and species-specific
acoustic monitoring. Most of the current methods are either manual or
semiautomatic, requireing labor-intensive processing/tuning.

In response to the need for automatic, tuning-free bird presentce detection
algorithm, Bird Audio Detection Challenge was created, hosted by Queen Mary
University of London and other universities. Three datasets were released
as part of the challenge, and teams across the world compete for the best
performance on the held-out dataset.

In section 2, we will discuss some of the competition details, including
the datasets used, task description and the evaluation metrics. Section 3
will present three methods for bird acoustic detection, two of which are
our submission to the challenge, and the third one being the best
performing method submitted by Thomas Grill. Section 4 will discuss the
evaluation results of the three methods, and Section 5 will present future
possible work based off of this challenge.

%------------------------------------------------------------------------
\section{Competition Details}

The Bird Audio Detection Challenge is hosted by Queen Mary University of
London, in collaboration with the IEEE Signal Processing Society.

%-------------------------------------------------------------------------
\subsection{Task Description}

For the challenge, teams have to design a system that takes a short audio
clip, and returns a binary label denoting the presence/absence of bird
sound of any kind. Teams are also encouraged to output a continuous,
probability score in the range of 0 to 1 for evaluation purposes.

%-------------------------------------------------------------------------
\subsection{Datasets}

Three datasets are provided as part of the challenge.

\textbf{freefield1010:} 7000 excerpts from field recordings around the
world, gathered by the FreeSound project.

\textbf{Warblr:} Crowdsourced project with many smartphone recordings from
around the United Kingdom. Very diverse in sound categories, even including
human bird imitations.

\textbf{Chernobyl:} Dataset created as part of Dr Wood's research into the
long-term effects of the Chernobyl accident on local ecology. Approximately
10,000 hours to date

Each dataset contains audio clips of roughly 10 seconds. For the challenge,
all datasets' audio files are provided, but only \textbf{freefield} and
\textbf{Warblr}'s labels are given. Held-out dataset consists primarily of
the \textbf{Chernobyl} dataset, along with some from the other two.

%-------------------------------------------------------------------------
\subsection{Evaluation Metrics}

To evaluate the performance of a model, we use the standard "Area Under the
ROC Curve" method, also known as the AUC score.

%-------------------------------------------------------------------------
\section{Bird Acoustic Detection}

Here we introduce three methods for bird acoustic detection, two of which
are proposed by us, and the third one being the best performing method by
Thomas Grill of University of Music and Performing Arts, Vienna.

%-------------------------------------------------------------------------
\subsection{Data preparation and augmentation}

All audio clips are 10 seconds, mono at 44100 Hz, which gives us 441000
frames of audio. We divide all frames by 32768 to convert values of
individual frames to between -1 and 1, and randomly crop out 400000 frames
as input to our network.

We also deployed data augmentation techniques. Given one base sound clip,
positive or negative, and another negative sound clip, we concatenate them
along the y-axis, with the base clip weighted by a factor \(r\) between 0.25
and 1, and the negative sound clip weighted by a factor \(1-r\). So we
have:
\begin{equation}
	x_{augmented}=r*x_{signal} + (1-r)*x_{negative}
\end{equation}
If \(r\) equals to 1, then there would be no augmentation, and if \(r\)
equals to 0.25, only 25\% of the base sound clip will be preserved. We
prevent \(r\) from going below 0.25 to ensure bird sound is still
detectable after augmentation.

Label for the augmented clip will remain the same as the base clip, as the
augmentation is an \(|\) process, preserving the original label when \(|\)
with a negative label.

%-------------------------------------------------------------------------
\subsection{Method 1}

Our first method resembles basic characteristics of SoundNet and residual
net. The network is fully convolutional, with one-dimensional convolutions
followed by nonlinearities (ReLU) layers. This network is suitable for
audio signals, as it is invariant to translation and allows stacking of
convolutional layers for detecting high-level features.

Network also has shortcuts from earlier layers of the network to later
stages, allowing easier gradient back-propagation to early layers of the
network. See Table 1 for network details.

%------------------------------------------------------------------------

\subsection{Method 2}

Our second method utilizes a series of pooling and convolutional layers to
extract sets of features from the input audio at different resolutions, and
concatenate them together before some final convolutional layers.

Network can be divided into three stages. In the first stage, we first
create two copies of the input, one reshaped to 100000 by 4, and the other
reshaped to 200000 by 2. \(x_{100000*4}\) is then passed to a convolutional
layer with kernel size \([3,4]\) while \(x_{200000*2}\) is passed to a
series of 5 pooling layers, followed by a convolutional layer at the end of
each pooling layer. Note here that the input of one pooling layer is the
output of the previous pooling layer, the outputs of the convolutional
layers serve as features extracted at audio of different resolution.

In the second stage, we combined the outputs of the 6 convolutional layers
in Stage 1 into one feature vector. First, we find the maximum for each
convolutional output along the thrid dimension. We denote these as the
\textit{energies} for each convolutional layer. We then concatenate them
with their corresponding convolutional output, run each concatenated
feature vector through a pooling layer of specified size, and concatenate
all 6 feature vectors into one along the third dimension. Due to different
lengths of output for the 6 convolutional layers, we pass them through
pooling layers of different sizes to ensure that they are of the same
dimension for concatenation.

In the final stage, we pass the output from Stage 2 through 7 additional
convolutional layers, with a \(max()\) operation between the 4th and 5th
layer. Final output is then flattened into size (1,2), ready to be used in
a softmax classification loss function.

%------------------------------------------------------------------------

\subsection{Method 3}

This method is the winning algorithm, submitted by Thomas Grill.

%------------------------------------------------------------------------
\section{Final copy}

You must include your signed IEEE copyright release form when you submit
your finished paper. We MUST have this form before your paper can be
published in the proceedings.

{\small \bibliographystyle{ieee} \bibliography{biblio} }

\end{document}

### 2017-03-15 src/dataset.py
Fixed code so network v8 works

@@ -8,7 +8,7 @@ import os
 import glob
 import ops
 
-d = 200000 # number of audio samples for learning
+d = 400000 # number of audio samples for learning
 
 basedir = '../data/'
 

### 2017-03-15 src/network.py

@@ -296,6 +296,7 @@ def network_v8(net, is_training=True, activation_fn=tf.nn.relu,
 
     with slim.arg_scope(network_arg_scope(is_training=is_training,
         activation_fn=activation_fn)):
+        import ipdb; ipdb.set_trace()
 
         net_x4 = tf.reshape(net,(-1,100000,4,1))
         net_x2 = tf.reshape(net,(-1,200000,2,1))
@@ -322,7 +323,7 @@ def network_v8(net, is_training=True, activation_fn=tf.nn.relu,
 
         print(energies)
 
-        filts_sup = [tf.concat_v2((a,b),3) for a,b in zip(filts,energies)]
+        filts_sup = [tf.concat(3,(a,b)) for a,b in zip(filts,energies)]
         
         print(filts_sup)
 
@@ -332,7 +333,7 @@ def network_v8(net, is_training=True, activation_fn=tf.nn.relu,
 
         print(pyr_inv)
         
-        net = tf.concat_v2(pyr_inv,3)
+        net = tf.concat(3,pyr_inv)
         net = slim.batch_norm(net)
 
         net = slim.conv2d(net,64,[9,1],stride=(5,1))

### 2017-03-17 paper/main.tex
Added figures and finished most of writing

@@
 \usepackage{graphicx} \usepackage{amsmath} \usepackage{amssymb}
+
+\graphicspath{ {figures/} }
+
+\newenvironment{alt_enumerate}
+{ \begin{enumerate}
+		\setlength{\itemsep}{1pt}
+		\setlength{\parskip}{1pt}
+		\setlength{\parsep}{1pt} }
+	{ \end{itemize} }
 
@@
 %%%%%%%%% TITLE
-\title{\LaTeX\ Author Guidelines for CVPR Proceedings}
+\title{Bird Audio Detection Challenge}
 
@@
 %%%%%%%%% ABSTRACT
-\begin{abstract} TODO: Make Abstract last \end{abstract}
+\begin{abstract}
+
+	Acoustic detection of birds in the wild is useful in many biological
+	scenarios, including estimation of bird population and species-specific
+	acoustic monitoring. Most of the current methods are either manual or
+	semiautomatic, requiring labor-intensive processing/tuning. We hereby
+	present three alternative methods, as part of the Bird Audio Detection
+	Challenge, hosted by Queen Mary University of London. Two methods are of
+	our own submission, with the third one as the winning solution submitted
+	by Thomas Grill. All methods utilize deep learning for automatic,
+	tuning-free bird presence detection.
+
+\end{abstract}
 
@@
 %-------------------------------------------------------------------------
-\subsection{Data preparation and augmentation}
+\subsection{Data preparation}
 
@@
 as input to our network.
-
-We also deployed data augmentation techniques. Given one base sound clip,
-positive or negative, and another negative sound clip, we concatenate them
-along the y-axis, with the base clip weighted by a factor \(r\) between 0.25
-and 1, and the negative sound clip weighted by a factor \(1-r\). So we
-have:
-\begin{equation}
-	x_{augmented}=r*x_{signal} + (1-r)*x_{negative}
-\end{equation}
-If \(r\) equals to 1, then there would be no augmentation, and if \(r\)
-equals to 0.25, only 25\% of the base sound clip will be preserved. We
-prevent \(r\) from going below 0.25 to ensure bird sound is still
-detectable after augmentation.
-
-Label for the augmented clip will remain the same as the base clip, as the
-augmentation is an \(|\) process, preserving the original label when \(|\)
-with a negative label.
 
@@
 stages, allowing easier gradient back-propagation to early layers of the
-network. See Table 1 for network details.
+network. See Figure 1 for network details.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.5\textwidth]{v5_diagram}
+	\caption{Network v5 Architecture. Content with box refers to kernel size,
+	number of filters, and stride size}
+\end{figure}
 
@@
 
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{v8_diagram}
+	\caption{Network v8 Architecture. Content of convolutional blocks refer
+	to kernel size, number of filters, and stride size. Content of avgpool
+	blocks refer to kernel size and stride size.}
+\end{figure*}
+
 %------------------------------------------------------------------------
-
 \subsection{Method 3}
 
-This method is the winning algorithm, submitted by Thomas Grill.
+% TODO: Explain that spectrograms were used during training
+
+This method is the winning algorithm of the challenge, submitted by Thomas
+Grill. The neural network used consists of 4 convolutional layers with
+pooling, and 2 dense layers. Multiple signal pre-processing techniques were
+used, including silence/noise trimming, and a different data augmentation
+technique involving shifting the spectra in time and frequency.
+Pseudo-labeling and bagging techniques are also deployed, which we will
+discuss in section 4.
 
 %------------------------------------------------------------------------
-\section{Final copy}
+\section{Experiments}
 
-You must include your signed IEEE copyright release form when you submit
-your finished paper. We MUST have this form before your paper can be
-published in the proceedings.
+During our development and training of the network, we experimented with
+various pre- and postprocessing techniques, as well as some training
+strategies.
+
+%------------------------------------------------------------------------
+\subsection{Implementation Details}
+
+The given labeled data from both dataset is split into 10 folds. Within
+each fold, number of positive and negative samples are balanced. During
+training, number of samples loaded from each dataset are also balanced, to
+ensure no bias.
+
+For in-house testing, 9 folds are used for training and 1 folf for testing.
+For submission, all 10 folds are used for training.
+
+For our method 1 and 2, we implemented them in python with Tensorflow
+machine learning library. We trained from scratch, with learning rate of
+0.03, learning rate decay by a factor of 0.1 every 40000 iterations, and a
+total of 300000 iterations.
+
+%------------------------------------------------------------------------
+\subsection{Data Augmentation}
+
+We deployed data augmentation as our only pre-processing technique. Given
+one base sound clip, positive or negative, and another negative sound clip,
+we concatenate them along the y-axis, with the base clip weighted by a
+factor \(r\) between 0.25 and 1, and the negative sound clip weighted by a
+factor \(1-r\). So we have:
+\begin{equation}
+	x_{augmented}=r*x_{signal} + (1-r)*x_{negative}
+\end{equation}
+If \(r\) equals to 1, then there would be no augmentation, and if \(r\)
+equals to 0.25, only 25\% of the base sound clip will be preserved. We
+prevent \(r\) from going below 0.25 to ensure bird sound is still
+detectable after augmentation.
+
+Label for the augmented clip will remain the same as the base clip, as the
+augmentation is an \(|\) process, preserving the original label when \(|\)
+with a negative label.
+
+%------------------------------------------------------------------------
+\subsection{Reducing capacity}
+
+One main difficulty we faced is the small size of training data. With no
+regularization, our networks consistently over-fit to the training data,
+perform near perfect on training and our own hold-out test data, but
+perform poorly on the challenge data. To compensate for the lack of data,
+we reduce the capacity of our models by reducing the number of filters for
+each convolutional layer. We do so until the network cannot perform
+perfectly on the training and testing data.
+
+%------------------------------------------------------------------------
+\subsection{Prediction Bagging}
+
+This post-processing technique averages prediction scores from many models.
+Due to the random cropping of audio signals during the data loading phase,
+different evaluation of the same audio clip will yield slightly different
+results. Therefore we predict 100 sets of scores for the held-out challenge
+dataset, and average them to obtain one set of scores that was submitted.
+
+%------------------------------------------------------------------------
+\subsection{Domain Confusion}
+
+% TODO Consult with Dr. Jacobs on whether this should be mentioned or not
+
+This training strategy addresses the issue of dataset bias. It is apparent
+that the three given datasets have disjoint domains. In addition, within
+each dataset, data is skewed either towards mostly positive samples or
+mostly negative ones. During training, it is very likely that the network
+will cheat by identifying which dataset the clip is from, and make its
+prediction based off of that information. In response, we take the audio
+features near the end of the network, pass it through another convolutional
+layer, and...
+
+%------------------------------------------------------------------------
+\subsection{Pseudo Labeling}
+
+This technique is used by many contestants and attempted by us. The
+assumption is that high-confidence predictions on the held-out challenge
+dataset are accurate enough to be used as extra data for retraining or
+fine-tuning. Basic procedures are:
+
+\begin{alt_enumerate}
+	\item{Train a model with only labeled data}
+	\item{Predict scores of the held-out challenge dataset}
+	\item{Find threshold where within it, predictions are accurate enough to
+		be used as extra labels}
+	\item{Add all predictions within threshold, and retrain or fine-tune
+		network}
+\end{alt_enumerate}
+
+%------------------------------------------------------------------------
+\section{Results}
+
+Below are the final AUC scores for each of the 3 methods:\\
+
+\begin{tabular}{ |p{2cm}|p{2cm}|p{2cm}|p{2cm}| }
+	\hline
+	\multicolumn{2}{|c|}{AUC Scores}\\
+	\hline
+	Method & AUC Score\\
+	\hline
+	Method 1 & 77.9\\
+	Method 2 & ??.?\\
+	Method 3 & 88.7\\
+	\hline
+\end{tabular}
+
+%------------------------------------------------------------------------
+\section{Future Work}
+
+For method 1 and 2, no signal pre-processing techniques were deployed,
+resulting in very noisy inputs to our network. It will be interesting to
+find out if pre-processing will aid the prediction accuracy of our network.
+
+%------------------------------------------------------------------------
+\section{Conclusion}
+
+Bird acoustic detection is an important task for many biological
+applications, and the challenge has advanced this field in many ways. We
+presented three methods, first two with a focus on unique network
+architecture, and the third one as the embodiment of many years of machine
+listening research by Thomas Grill. Many topics are still open for
+research, like bird species identification and detection.
+
+%------------------------------------------------------------------------
 

### 2017-03-17 paper/main.tex
Polished paper a bit

@@
 		\setlength{\parsep}{1pt} }
-	{ \end{itemize} }
+	{ \end{enumerate} }
 
@@
 University of Kentucky\\ 
-{\tt\small
-weilian.song@uky.edu}
+{\tt\small weilian.song@uky.edu}
 % For a paper whose authors are all at the same institution, omit the
@@
 
+% TODO: Add Tawfiq in without messing up format
+
 \maketitle
@@
 	scenarios, including estimation of bird population and species-specific
-	acoustic monitoring. Most of the current methods are either manual or
-	semiautomatic, requiring labor-intensive processing/tuning. We hereby
+	acoustic monitoring. Most of current methods are either manual or
+	semi-automatic, requiring labor-intensive processing/tuning. We hereby
 	present three alternative methods, as part of the Bird Audio Detection
-	Challenge, hosted by Queen Mary University of London. Two methods are of
-	our own submission, with the third one as the winning solution submitted
-	by Thomas Grill. All methods utilize deep learning for automatic,
+	Challenge, hosted by Queen Mary University of London. Two methods are our
+	own submission, with the third one as the winning solution submitted by
+	Thomas Grill. All methods utilize deep neural networks for automatic,
 	tuning-free bird presence detection.
@@
 
-Acoustic detection of birds in the wild is useful in many biological
-scenarios, including estimation of bird population and species-specific
-acoustic monitoring. Most of the current methods are either manual or
-semiautomatic, requireing labor-intensive processing/tuning.
+Much biological research relies on bird acoustic detection in audio signals,
+as indicated by \cite{bad}. Over the recent years, a large amount of audio
+data is recorded, but the bird detection methods are often manual or
+semi-automatic, requiring intensive labor or tuning.
 
 In response to the need for automatic, tuning-free bird presentce detection
-algorithm, Bird Audio Detection Challenge was created, hosted by Queen Mary
-University of London and other universities. Three datasets were released
-as part of the challenge, and teams across the world compete for the best
-performance on the held-out dataset.
+algorithm, \textit{Bird Audio Detection Challenge} was created, hosted by
+Queen Mary University of London and other universities. Three datasets were
+released as part of the challenge, and teams across the world compete for
+the best performance on the task of bird audio detection.
 
-In section 2, we will discuss some of the competition details, including
-the datasets used, task description and the evaluation metrics. Section 3
-will present three methods for bird acoustic detection, two of which are
-our submission to the challenge, and the third one being the best
-performing method submitted by Thomas Grill. Section 4 will discuss the
-evaluation results of the three methods, and Section 5 will present future
-possible work based off of this challenge.
+In section 2, we discuss some of the competition details, including the
+datasets used, task description and the evaluation metrics. Section 3
+presents three methods for bird acoustic detection, two of which are our
+submission to the challenge, and the third one being the best performing
+method submitted by Thomas Grill. Section 4 discusses some experiments and
+techniques used to improve the peformance. Section 5 displays the
+evaluation results of the three methods, and section 6 presents future
+possible work with this challenge.
 
@@
 clip, and returns a binary label denoting the presence/absence of bird
-sound of any kind. Teams are also encouraged to output a continuous,
+sound of any species. Teams are also encouraged to output a continuous,
 probability score in the range of 0 to 1 for evaluation purposes.
@@
 
-\textbf{freefield1010:} 7000 excerpts from field recordings around the
+% TODO: Should FreeSound be italicized?
+
+\textbf{freefield1010:} 7,000 excerpts from field recordings around the
 world, gathered by the FreeSound project.
 
-\textbf{Warblr:} Crowdsourced project with many smartphone recordings from
+\textbf{Warblr:} Crowd-sourced project with many smartphone recordings from
 around the United Kingdom. Very diverse in sound categories, even including
@@
 
-\textbf{Chernobyl:} Dataset created as part of Dr Wood's research into the
-long-term effects of the Chernobyl accident on local ecology. Approximately
-10,000 hours to date
+% TODO: Should link to project be included?
+
+\textbf{Chernobyl:} Part of TREE (Transfer-Exposure-Effects) research
+project. Consists of audio in the Chernobyl Exclusion Zone.
 
@@
 \textbf{Warblr}'s labels are given. Held-out dataset consists primarily of
-the \textbf{Chernobyl} dataset, along with some from the other two.
+the \textbf{Chernobyl} dataset, along with some from \textbf{freefield} and
+\textbf{Warblr}.
 
@@
 %-------------------------------------------------------------------------
-\section{Bird Acoustic Detection}
+\section{Methodology}
 
@@
 
-All audio clips are 10 seconds, mono at 44100 Hz, which gives us 441000
-frames of audio. We divide all frames by 32768 to convert values of
-individual frames to between -1 and 1, and randomly crop out 400000 frames
-as input to our network.
+All individual audio clips are 10 seconds, mono at 44,100 Hz, which gives us
+441,000 frames of audio. We divide all frames by 32,768 to convert values of
+individual frames to between -1 and 1, and randomly crop out 400,000
+continuous frames, resulting in a one-dimensional vector with length 400,000
+as our input to the network.
 
@@
 
-Our first method resembles basic characteristics of SoundNet and residual
-net. The network is fully convolutional, with one-dimensional convolutions
-followed by nonlinearities (ReLU) layers. This network is suitable for
-audio signals, as it is invariant to translation and allows stacking of
+Our first method resembles basic characteristics of SoundNet by Aytar et
+al.\cite{soundnet} and Residual Net by He et al.\cite{residual_net}. The
+network is fully convolutional, with one-dimensional convolutions followed
+by nonlinearities (ReLU) layers.  This network is suitable for audio
+signals, as it is invariant to translation and allows stacking of
 convolutional layers for detecting high-level features.
@@
 	\includegraphics[width=0.5\textwidth]{v5_diagram}
-	\caption{Network v5 Architecture. Content with box refers to kernel size,
-	number of filters, and stride size}
+	\caption{Network v5 Architecture. Content within convolutional blocks
+	refers to kernel size, number of filters, and stride size}
 \end{figure}
@@
 Network can be divided into three stages. In the first stage, we first
-create two copies of the input, one reshaped to 100000 by 4, and the other
-reshaped to 200000 by 2. \(x_{100000*4}\) is then passed to a convolutional
-layer with kernel size \([3,4]\) while \(x_{200000*2}\) is passed to a
-series of 5 pooling layers, followed by a convolutional layer at the end of
-each pooling layer. Note here that the input of one pooling layer is the
-output of the previous pooling layer, the outputs of the convolutional
-layers serve as features extracted at audio of different resolution.
+create two copies of the input, one reshaped to 100,000 by 4, and the other
+reshaped to 200,000 by 2. \(x_{100,000*4}\) is then passed through a single
+convolutional layer while \(x_{200,000*2}\) is passed through a series of 5
+pooling layers, followed by a convolutional layer at the end of each
+pooling layer. Note here that the input of one pooling layer is the output
+of the previous pooling layer, the outputs of the convolutional layers
+serve as features extracted at input of different resolution.
 
@@
 in Stage 1 into one feature vector. First, we find the maximum for each
-convolutional output along the thrid dimension. We denote these as the
-\textit{energies} for each convolutional layer. We then concatenate them
-with their corresponding convolutional output, run each concatenated
-feature vector through a pooling layer of specified size, and concatenate
-all 6 feature vectors into one along the third dimension. Due to different
-lengths of output for the 6 convolutional layers, we pass them through
-pooling layers of different sizes to ensure that they are of the same
-dimension for concatenation.
+convolutional output. We denote them as the energy for each convolutional
+layer. We then concatenate them with their corresponding convolutional
+output, pass each concatenated feature vector through a pooling layer of
+specified size, and concatenate all 6 feature vectors into one. Due to
+different lengths of output for the 6 convolutional layers in stage 1, we
+pass them through pooling layers of different sizes to ensure that they are
+of the same dimension for concatenation.
 
-In the final stage, we pass the output from Stage 2 through 7 additional
-convolutional layers, with a \(max()\) operation between the 4th and 5th
-layer. Final output is then flattened into size (1,2), ready to be used in
-a softmax classification loss function.
+In the final stage, we pass the output from Stage 2 through a batch
+normalization layer, followed by 7 additional convolutional layers, with a
+\(max()\) operation between the 4th and 5th layer. Final output is then
+flattened into size (1,2), ready to be used in a softmax classification
+loss function.
 
@@
 
-This method is the winning algorithm of the challenge, submitted by Thomas
-Grill. The neural network used consists of 4 convolutional layers with
-pooling, and 2 dense layers. Multiple signal pre-processing techniques were
-used, including silence/noise trimming, and a different data augmentation
-technique involving shifting the spectra in time and frequency.
-Pseudo-labeling and bagging techniques are also deployed, which we will
-discuss in section 4.
+This method is the winning algorithm to the challenge, submitted by Thomas
+Grill. The neural network uses processed spectrograms of audio clips as
+input, instead of a one-dimensional vector of signals.The network consists
+of 4 convolutional layers with pooling, and 2 dense layers. Multiple signal
+pre-processing techniques are used, including silence/noise trimming, and a
+different data augmentation technique involving shifting the spectrograms
+in time and frequency.  Pseudo-labeling and bagging techniques are also
+deployed, which we will discuss in section 4.
 
@@
 
-During our development and training of the network, we experimented with
-various pre- and postprocessing techniques, as well as some training
-strategies.
+During our development, we experimented with various pre- and
+postprocessing techniques, as well as some training strategies.
 
@@
 
-The given labeled data from both dataset is split into 10 folds. Within
+The given labeled data from both datasets is split into 10 folds. Within
 each fold, number of positive and negative samples are balanced. During
-training, number of samples loaded from each dataset are also balanced, to
+training, number of samples loaded from each datasets are also balanced, to
 ensure no bias.
 
-For in-house testing, 9 folds are used for training and 1 folf for testing.
+For in-house testing, 9 folds are used for training and 1 fold for testing.
 For submission, all 10 folds are used for training.
 
-For our method 1 and 2, we implemented them in python with Tensorflow
-machine learning library. We trained from scratch, with learning rate of
-0.03, learning rate decay by a factor of 0.1 every 40000 iterations, and a
-total of 300000 iterations.
+For method 1 and 2, we implemented them in python with Tensorflow machine
+learning library. We trained from scratch, with learning rate of 0.03,
+learning rate decay by a factor of 0.1 every 40,000 iterations, and a total
+of 300,000 iterations.
+
+For method 3, code base is available online at...
+
+% TODO: How to include code base link?
 
@@
 
-We deployed data augmentation as our only pre-processing technique. Given
-one base sound clip, positive or negative, and another negative sound clip,
-we concatenate them along the y-axis, with the base clip weighted by a
-factor \(r\) between 0.25 and 1, and the negative sound clip weighted by a
-factor \(1-r\). So we have:
+We deploy data augmentation as our only pre-processing technique for method
+1 and 2. Given one base sound clip, positive or negative (bird
+present/absent), and another negative sound clip, we concatenate them along
+the y-axis, with the base clip weighted by a factor \(r\) between 0.25 and
+1, and the negative sound clip weighted by a factor \(1-r\). So we have:
 \begin{equation}
-	x_{augmented}=r*x_{signal} + (1-r)*x_{negative}
+	x_{augmented}=r*x_{base} + (1-r)*x_{negative}
 \end{equation}
@@
 equals to 0.25, only 25\% of the base sound clip will be preserved. We
-prevent \(r\) from going below 0.25 to ensure bird sound is still
-detectable after augmentation.
+sample \(r\) on a positively skewed interval (0.25,1) to ensure most bird
+sounds in augmented clips are still detectable.
 
 Label for the augmented clip will remain the same as the base clip, as the
-augmentation is an \(|\) process, preserving the original label when \(|\)
-with a negative label.
+augmentation is an \textit{or} process, preserving the original label when 
+\textit{or} with a negative label.
 
@@
 regularization, our networks consistently over-fit to the training data,
-perform near perfect on training and our own hold-out test data, but
-perform poorly on the challenge data. To compensate for the lack of data,
-we reduce the capacity of our models by reducing the number of filters for
-each convolutional layer. We do so until the network cannot perform
-perfectly on the training and testing data.
+perform near perfect on training and our own held-out test data, but
+perform poorly on the challenge data. To compensate this, we reduce the
+capacity of our models by reducing the number of filters for each
+convolutional layer. We do so until the network cannot perform perfectly on
+the training and testing data.
 
@@
 
-This post-processing technique averages prediction scores from many models.
-Due to the random cropping of audio signals during the data loading phase,
-different evaluation of the same audio clip will yield slightly different
-results. Therefore we predict 100 sets of scores for the held-out challenge
-dataset, and average them to obtain one set of scores that was submitted.
+This post-processing technique averages prediction scores from many
+sessions. Due to the random cropping of audio signals during the data
+loading phase, different evaluation of the same audio clip will yield
+slightly different results. Therefore we predict 100 sets of scores for the
+held-out challenge dataset, and average them to obtain one set of scores
+that was submitted.
+
+For method 3, instead of averaging sets of output scores from the same
+model, different models' outputs were averaged. Please refer to Grill's
+code base for more details.
 
@@
 This training strategy addresses the issue of dataset bias. It is apparent
-that the three given datasets have disjoint domains. In addition, within
-each dataset, data is skewed either towards mostly positive samples or
-mostly negative ones. During training, it is very likely that the network
-will cheat by identifying which dataset the clip is from, and make its
-prediction based off of that information. In response, we take the audio
-features near the end of the network, pass it through another convolutional
-layer, and...
+that the three given datasets have disjoint sound domains. In addition,
+within each dataset, data is skewed either towards mostly positive samples
+or mostly negative ones. During training, it is very likely that the
+network will cheat by identifying which dataset the clip is from, and make
+its prediction based off of that information. In response, we take the
+audio features near the end of the network, pass it through another
+convolutional layer, and...
 
@@
 	\item{Predict scores of the held-out challenge dataset}
-	\item{Find threshold where within it, predictions are accurate enough to
-		be used as extra labels}
+	\item{Find threshold \(r\) such that predictions within range\\
+		\(0<r \cup (1-r)<1\) are accurate enough}
 	\item{Add all predictions within threshold, and retrain or fine-tune
@@
 \end{alt_enumerate}
+
+\(r\) can be obtained through manual spot-checking of the predictions. Too
+much time spend on finding the threshold will defeat the purpose of the
+challenge.
 
@@
 resulting in very noisy inputs to our network. It will be interesting to
-find out if pre-processing will aid the prediction accuracy of our network.
+find out if pre-processing will aid the prediction accuracy of our two
+networks.
 
@@
 Bird acoustic detection is an important task for many biological
-applications, and the challenge has advanced this field in many ways. We
-presented three methods, first two with a focus on unique network
+applications, but the massive amount of data available makes it difficult
+to extract any useful information from it. The Bird Audio Detection
+challenge has advanced this field in many ways, and many excellent
+automatic, tuning-free bird acoustic detection algorithms are introduced.
+We presented three methods, first two with a focus on unique network
 architecture, and the third one as the embodiment of many years of machine
-listening research by Thomas Grill. Many topics are still open for
-research, like bird species identification and detection.
+listening research.  Many topics are still open for research, including
+many tasks that already have a solution, but only achieved through manual
+or labor intensive.
 

### 2017-03-18 paper/main.tex
Added entry to table
More edits to paper

@@
 
-\author{Weilian Song\\ 
+\author{
+	Weilian Song, Tawfiq Salem, Nathan Jacobs\\ 
 	University of Kentucky\\ 
-{\tt\small weilian.song@uky.edu}
-% For a paper whose authors are all at the same institution, omit the
-% following lines up until the closing ``}''.  Additional authors and
-% addresses can be added with ``\and'', just like the second author.  To
-% save space, use either the email address or home page, not both
-\and Nathan Jacobs\\ 
-University of Kentucky\\ 
-{\tt\small jacobs@cs.uky.edu}
+	{\tt\small weilian.song@uky.edu, tawfiq.salem@uky.edu, jacobs@cs.uky.edu}
 }
@@
 	Thomas Grill. All methods utilize deep neural networks for automatic,
-	tuning-free bird presence detection.
+	tuning-free bird presence detection, with the first two methods
+	emphasizing network architectures, and the third one utilizing multiple
+	signal preprocessing techniques.
 
@@
 evaluation results of the three methods, and section 6 presents future
-possible work with this challenge.
+possible work from this challenge.
 
@@
 clip, and returns a binary label denoting the presence/absence of bird
-sound of any species. Teams are also encouraged to output a continuous,
+sound of any species. Teams are also encouraged to output a floating-point,
 probability score in the range of 0 to 1 for evaluation purposes.
@@
 \textbf{Chernobyl:} Part of TREE (Transfer-Exposure-Effects) research
-project. Consists of audio in the Chernobyl Exclusion Zone.
+project. Consists of ambient audio in the Chernobyl Exclusion Zone.
 
@@
 all datasets' audio files are provided, but only \textbf{freefield} and
-\textbf{Warblr}'s labels are given. Held-out dataset consists primarily of
-the \textbf{Chernobyl} dataset, along with some from \textbf{freefield} and
-\textbf{Warblr}.
+\textbf{Warblr}'s labels are given. Held-out challenge dataset consists
+primarily of the \textbf{Chernobyl} dataset, along with some from
+\textbf{freefield} and \textbf{Warblr}.
 
@@
 network is fully convolutional, with one-dimensional convolutions followed
-by nonlinearities (ReLU) layers.  This network is suitable for audio
+by non-linearities (ReLU) layers.  This network is suitable for audio
 signals, as it is invariant to translation and allows stacking of
@@
 Our second method utilizes a series of pooling and convolutional layers to
-extract sets of features from the input audio at different resolutions, and
-concatenate them together before some final convolutional layers.
+extract features from the input audio at different temporal resolutions,
+and concatenate them together before some final convolutional layers.
 
@@
 convolutional layer while \(x_{200,000*2}\) is passed through a series of 5
-pooling layers, followed by a convolutional layer at the end of each
-pooling layer. Note here that the input of one pooling layer is the output
-of the previous pooling layer, the outputs of the convolutional layers
-serve as features extracted at input of different resolution.
+averaging pooling layers, followed by a convolutional layer at the end of
+each pooling layer. Note here that the input of one pooling layer is the
+output of the previous pooling layer, the outputs of the convolutional
+layers serve as features extracted at input of different temporal
+resolution.
 
@@
 output, pass each concatenated feature vector through a pooling layer of
-specified size, and concatenate all 6 feature vectors into one. Due to
-different lengths of output for the 6 convolutional layers in stage 1, we
-pass them through pooling layers of different sizes to ensure that they are
-of the same dimension for concatenation.
+specified size, and concatenate all 6 pooled feature vectors into one. Due
+to different lengths of output from the 6 convolutional layers in stage 1,
+we pass them through pooling layers of different sizes to ensure that they
+are of the same dimension for concatenation.
 
@@
 The given labeled data from both datasets is split into 10 folds. Within
-each fold, number of positive and negative samples are balanced. During
-training, number of samples loaded from each datasets are also balanced, to
-ensure no bias.
+each fold, number of positive(bird present) and negative(bird absent)
+samples are balanced. During training, number of samples loaded from each
+datasets are also balanced, to ensure no bias.
 
@@
 
-For method 3, code base is available online at...
+For method 3, a brief description and code base is available online
+at\cite{winning_blog}
 
@@
 equals to 0.25, only 25\% of the base sound clip will be preserved. We
-sample \(r\) on a positively skewed interval (0.25,1) to ensure most bird
-sounds in augmented clips are still detectable.
+sample \(r\) on a positively skewed interval (0.25,1) to ensure bird sounds
+in positive labels are not diminished.
 
@@
 slightly different results. Therefore we predict 100 sets of scores for the
-held-out challenge dataset, and average them to obtain one set of scores
-that was submitted.
+held-out challenge dataset, and average them to capture the true scores for
+submission.
 
@@
 model, different models' outputs were averaged. Please refer to Grill's
-code base for more details.
+code base\cite{winning_blog} for more details.
 
@@
 	Method 1 & 77.9\\
-	Method 2 & ??.?\\
+	Method 2 & 78.8\\
 	Method 3 & 88.7\\

### 2017-03-20 paper/main.tex
Fixed many grammatical mistakes

@@
 	scenarios, including estimation of bird population and species-specific
-	acoustic monitoring. Most of current methods are either manual or
+	acoustic monitoring. Most current methods are either manual or
 	semi-automatic, requiring labor-intensive processing/tuning. We hereby
 	present three alternative methods, as part of the Bird Audio Detection
-	Challenge, hosted by Queen Mary University of London. Two methods are our
-	own submission, with the third one as the winning solution submitted by
-	Thomas Grill. All methods utilize deep neural networks for automatic,
-	tuning-free bird presence detection, with the first two methods
-	emphasizing network architectures, and the third one utilizing multiple
+	Challenge, hosted by Queen Mary University of London. The first two
+	methods are of our own submission, the third being the winning solution
+	submitted by Thomas Grill. Each of said methods utilize deep neural
+	networks for automatic, tuning-free bird presence detection: the first
+	two emphasize network architectures, while the third utilizes multiple
 	signal preprocessing techniques.
@@
 
-In response to the need for automatic, tuning-free bird presentce detection
-algorithm, \textit{Bird Audio Detection Challenge} was created, hosted by
-Queen Mary University of London and other universities. Three datasets were
-released as part of the challenge, and teams across the world compete for
-the best performance on the task of bird audio detection.
+In response to the need for an automatic, tuning-free bird presence
+detection algorithm, \textit{Bird Audio Detection Challenge} was created,
+hosted by Queen Mary University of London and other universities. Three
+datasets were released as part of the challenge, and teams across the world
+compete for the best performance on the task of bird audio detection.
 
@@
 evaluation results of the three methods, and section 6 presents future
-possible work from this challenge.
+possible work that may result from this challenge.
 
@@
 For the challenge, teams have to design a system that takes a short audio
-clip, and returns a binary label denoting the presence/absence of bird
-sound of any species. Teams are also encouraged to output a floating-point,
+clip, and returns a binary label denoting the presence/absence of any bird
+species' sound. Teams are also encouraged to output a floating-point,
 probability score in the range of 0 to 1 for evaluation purposes.
@@
 
-\textbf{Warblr:} Crowd-sourced project with many smartphone recordings from
-around the United Kingdom. Very diverse in sound categories, even including
-human bird imitations.
+\textbf{Warblr:} Crowd-sourced project with many smartphone recordings.
+Very diverse in sound categories, even including human bird imitations.
 
@@
 
+% TODO: Weird spacing issue in first and second sentence
+
 Our first method resembles basic characteristics of SoundNet by Aytar et
@@
 network is fully convolutional, with one-dimensional convolutions followed
-by non-linearities (ReLU) layers.  This network is suitable for audio
+by non-linear (ReLU) layers. This network is suitable for audio
 signals, as it is invariant to translation and allows stacking of
@@
 
-Network also has shortcuts from earlier layers of the network to later
+This network also has shortcuts from earlier layers of the network to later
 stages, allowing easier gradient back-propagation to early layers of the
@@
 extract features from the input audio at different temporal resolutions,
-and concatenate them together before some final convolutional layers.
+and concatenate them before some final convolutional layers.
 
-Network can be divided into three stages. In the first stage, we first
+This network can be divided into three stages. In the first stage, we first
 create two copies of the input, one reshaped to 100,000 by 4, and the other
@@
 each pooling layer. Note here that the input of one pooling layer is the
-output of the previous pooling layer, the outputs of the convolutional
+output of the previous pooling layer; the outputs of the convolutional
 layers serve as features extracted at input of different temporal
@@
 
-In the second stage, we combined the outputs of the 6 convolutional layers
-in Stage 1 into one feature vector. First, we find the maximum for each
+In the second stage, we combine the outputs of the 6 convolutional layers
+in stage 1 into one feature vector. First, we find the maximum for each
 convolutional output. We denote them as the energy for each convolutional
@@
 output, pass each concatenated feature vector through a pooling layer of
-specified size, and concatenate all 6 pooled feature vectors into one. Due
-to different lengths of output from the 6 convolutional layers in stage 1,
-we pass them through pooling layers of different sizes to ensure that they
-are of the same dimension for concatenation.
+specified size, and concatenate all 6 pooled feature vectors. Due to
+different lengths of output from the 6 convolutional layers in stage 1, we
+pass them through pooling layers of different sizes to ensure that they are
+of the same dimension for concatenation.
 
-In the final stage, we pass the output from Stage 2 through a batch
+In the final stage, we pass the output from stage 2 through a batch
 normalization layer, followed by 7 additional convolutional layers, with a
@@
 
-% TODO: Explain that spectrograms were used during training
-
 This method is the winning algorithm to the challenge, submitted by Thomas
 Grill. The neural network uses processed spectrograms of audio clips as
-input, instead of a one-dimensional vector of signals.The network consists
+input, instead of a one-dimensional vector of signals. The network consists
 of 4 convolutional layers with pooling, and 2 dense layers. Multiple signal
@@
 During our development, we experimented with various pre- and
-postprocessing techniques, as well as some training strategies.
+post-processing techniques, as well as some training strategies.
 
@@
 The given labeled data from both datasets is split into 10 folds. Within
-each fold, number of positive(bird present) and negative(bird absent)
-samples are balanced. During training, number of samples loaded from each
-datasets are also balanced, to ensure no bias.
+each fold, the number of positive(bird present) and negative(bird absent)
+samples are balanced. During training, the number of samples loaded from
+each datasets are also balanced, to ensure no bias.
 
@@
 For method 3, a brief description and code base is available online
-at\cite{winning_blog}
+at\cite{winning_blog}.
 
@@
 the y-axis, with the base clip weighted by a factor \(r\) between 0.25 and
-1, and the negative sound clip weighted by a factor \(1-r\). So we have:
+1, and the negative sound clip weighted by a factor \(1-r\). Thus we have:
 \begin{equation}
@@
 augmentation is an \textit{or} process, preserving the original label when 
-\textit{or} with a negative label.
+"\textit{or}-ed" with a negative label.
 
@@
 
-One main difficulty we faced is the small size of training data. With no
+One main difficulty we faced was the small size of training data. With no
 regularization, our networks consistently over-fit to the training data,
-perform near perfect on training and our own held-out test data, but
-perform poorly on the challenge data. To compensate this, we reduce the
-capacity of our models by reducing the number of filters for each
-convolutional layer. We do so until the network cannot perform perfectly on
-the training and testing data.
+performed nearly perfectly on training and our own held-out test data, but
+performed poorly on the challenge data. To compensate for this, we reduced
+the capacity of our models by reducing the number of filters for each
+convolutional layer. We did so until the network could not perform
+perfectly on the training and testing data.
 
@@
 loading phase, different evaluation of the same audio clip will yield
-slightly different results. Therefore we predict 100 sets of scores for the
-held-out challenge dataset, and average them to capture the true scores for
+slightly different results. Therefore we predicted 100 sets of scores for the
+held-out challenge dataset, and averaged them to capture the true scores for
 submission.
@@
 
-This technique is used by many contestants and attempted by us. The
+This technique is used by many contestants and was attempted by us. The
 assumption is that high-confidence predictions on the held-out challenge
@@
 \(r\) can be obtained through manual spot-checking of the predictions. Too
-much time spend on finding the threshold will defeat the purpose of the
+much time spent on finding the threshold will defeat the purpose of the
 challenge.
@@
 
-For method 1 and 2, no signal pre-processing techniques were deployed,
+For methods 1 and 2, no signal pre-processing techniques were deployed,
 resulting in very noisy inputs to our network. It will be interesting to
@@
 to extract any useful information from it. The Bird Audio Detection
-challenge has advanced this field in many ways, and many excellent
-automatic, tuning-free bird acoustic detection algorithms are introduced.
-We presented three methods, first two with a focus on unique network
-architecture, and the third one as the embodiment of many years of machine
+challenge has advanced this field greatly--many excellent automatic,
+tuning-free bird acoustic detection algorithms are introduced.  We
+presented three methods, the first two with a focus on unique network
+architectures, and the third one as the embodiment of many years of machine
 listening research.  Many topics are still open for research, including
-many tasks that already have a solution, but only achieved through manual
-or labor intensive.
+numerous tasks that already have a solution, but can only be achieved
+through manual or labor intensive annotation.
 
